---
title: "p8105_hw1_tp2806"
author: "Tejashree Prakash"
date: 2025-09-11
output: github_document
---

```{r setup, include=FALSE}
#Load libraries and data
library(moderndive)
library(ggplot2) #For ggplot scatter plot generation. 
data("early_january_weather")
```

```{r}
#Problem 1: Get size of the dataset
rows <- nrow(early_january_weather)
col <- ncol(early_january_weather)
mean_temp <- mean(early_january_weather$temp)
```

The early_january_weather dataset contains `r rows` rows and `r col` columns of hourly weather data collected during early January. The variables included in this dataset are `r ls(early_january_weather)`. The mean temperature from the values in this dataset is `r mean_temp`.

```{r}
#Problem 1: Make scatter plot. 
weather_scplot <- ggplot(early_january_weather, aes(x = time_hour, y=temp, colour = humid)) + geom_point()

#Save plot. 
ggsave("weather_scatterplot.png", plot = weather_scplot, width = 6, height = 4, dpi = 300)
```
The scatter plot demonstrates a general upward shift, when relating time by hour (x) against temperature (y). This means that, on average, as the time passes, the temperature increases. Additionally, the later times have a greater humidity level than compared to the earlier times in this dataset. 

```{r}
#Problem 2:
```

