---
title: "p8105_hw1_tp2806"
author: "Tejashree Prakash"
date: 2025-09-11
output: github_document
---

```{r setup, include=FALSE}
#Load libraries and data
library(moderndive)
library(ggplot2) #for ggplot scatter plot generation. 
library(tidyverse) #for dataframe manipulation 
data("early_january_weather")
```

```{r}
#Problem 1: Get size of the dataset
rows <- nrow(early_january_weather)
col <- ncol(early_january_weather)
mean_temp <- mean(early_january_weather$temp)
```

The early_january_weather dataset contains `r rows` rows and `r col` columns of hourly weather data collected during early January. The variables included in this dataset are `r ls(early_january_weather)`.

Important variables include:

- `temp`: temperature in °F, the mean temperature is `r mean_temp` °F.

- `humid`: relative humidity (%)


```{r}
#Problem 1: Make scatter plot. 
weather_scplot <- ggplot(early_january_weather, aes(x = time_hour, y=temp, colour = humid)) + geom_point()
weather_scplot

#Save plot. 
ggsave("weather_scatterplot.png", plot = weather_scplot, width = 6, height = 4, dpi = 300)
```
The scatter plot demonstrates a general upward shift, when relating time by hour (x) against temperature (y). This means that, on average, as the time passes, the temperature increases. Additionally, the later times have a greater humidity level than compared to the earlier times in this dataset. 

```{r}
#Problem 2: Create df. 
randomdf = tibble(
  numeric_samp = rnorm(10, mean = 1), #random sample of size 10, mean 1 
  greater_0 = numeric_samp > 0, #logical vector to indicate whether > 0 (TRUE/FALSE)
  char_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),#character vector of length 10
  factor_vector = factor(
    sample(c("low","medium","high"), size = 10, replace = TRUE), #create three levels 
    levels = c("low","medium","high")
))
print(randomdf)

#Perform means for different variables
mean_numeric <- mean(randomdf$numeric_samp)
mean_logic <- mean(randomdf$greater_0)
mean_char <- mean(randomdf$char_vector)
mean_factor <- mean(randomdf$factor_vector)
```
If I take the mean of my numeric vector of 10 random values, I get `r mean_numeric`. If I take the mean of the logic variable, I get `r mean_logic`, which is the proportion of values greater than 0. However, if I take the mean of the remaining two variables, the character and factor variables, I return NA. This is because these variables do not contain numeric values, and therefore I cannot obtain a mean. 


```{r results='hide', message=FALSE}
# Convert each variable type to numeric 
greater_0 <- as.numeric(randomdf$greater_0)
char_vector <- as.numeric(randomdf$char_vector)   
factor_vector <- as.numeric(randomdf$factor_vector)
```
If I force these variables to a numeric class, it mostly works. The logic variable still results in the same mean, which demonstrates that logic variables can work to produce a mean without being converted to numeric. For the factor variable, I am able to change it to numeric and then obtain a mean - which is the mean of underlying level, low = 1, medium = 2, high = 3 . However, a character variable cannot be converted to numeric unless I map it out to be numeric. For example, I'd have to state that a = 1, b = 2, etc. 
